\section{Conclusion and further work}

We took the algorithm of Tali Eilam-Tzoreff and significantly improved its running time to the upper bound of $\OO(|V|^6)$ for unweighted graphs and to $\OO(|V|^7)$ for the weighted graphs. 

We also made an extensive work of experimental evaluation of the original algorithm as well as its optimized versions. Taking into consideration the fact that the main theorem \ref{main_theorem} (that forms the basis for all of the algorithms considered here) is rather hard complicated by its formulation as well as by its technical proof (cf. \cite{ET}), its experimental evaluation if of certain benefit.

Let us also note the wide spectre of a further work in this area. First of all, one may improve the running time for unweighted or weighted undirected $2DSP$ even further, for example, by developing the linear algebra ideas that are founded in this work (which seems to be a perspective direction, considering the fact that what seems to be hardest part of the algorithm may indeed be optimized to the running time of $\OO(|V|^{3 + \omega})$). Then, there are open questions about the existence of a polynomial algorithm for the directed case of $2DSP$ (weighted or unit-length) and for the undirected version of $kDSP$ for $k \geq 3$ (in any formulation). 

As a final remark, note that all the given complexity upper bounds are expressed in terms of the number of vertices. It is possible that, by performing the more rigorous analysis, one may obtain a better upper limit in terms of the number of vertices as well as the number of edges in the graph, which will be more precise for the sparse graphs. 
